---
title: "Benchmarking mining commodity costs and their drivers"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Benchmarking mining commodity costs and their drivers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(MiningCostBench)
```

## What this package is

`MiningCostBench` bundles indicative cash/production cost ranges and a
structured, six-stage cost-driver breakdown for mining and mineral commodities,
across six categories. It is an orientation and teaching dataset, **not** a
substitute for a commercial cost-curve service.

## Categories and coverage

```{r}
mcb_categories()
table(commodity_costs$category)
```

## Looking up a commodity

```{r}
mcb_costs("copper")[, c("commodity", "cost_basis", "cost_text", "meaningful_metric")]
```

Matching is case-insensitive and falls back to partial matches:

```{r}
mcb_commodities("Precious Metals")
```

## The cost-driver breakdown

Every commodity is decomposed into six driver stages. `mcb_drivers()` returns
this in tidy long form, ready for `dplyr`/`ggplot2`:

```{r}
mcb_drivers("gold")
```

## By-product minerals

A structurally important subset of critical minerals are not mined on their own;
they are recovered as by-products and so have no standalone mining cost:

```{r}
mcb_byproducts()$commodity
```

## Cross-cutting drivers and sources

```{r}
head(mcb_cross_cutting(), 4)
mcb_sources("SOURCES")
```

## Caveats

Figures are indicative cash/production costs (not prices) and can vary 3-5x with
grade, method, depth, geography, currency and by-product credits. For many
critical minerals the **driver logic is more reliable than the dollar figure**.
Always verify against current company guidance or a commercial cost-curve
provider (Wood Mackenzie, CRU, S&P Global) before publication.
