---
title: "Benchmarking mining commodity costs and their drivers"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Benchmarking mining commodity costs and their drivers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(MiningCostBench)
```

## What this package is

`MiningCostBench` bundles indicative cash/production cost ranges and a
structured, six-stage cost-driver breakdown for mining and mineral commodities,
across six categories. It is an orientation and teaching dataset, **not** a
substitute for a commercial cost-curve service.

## Categories and coverage

```{r}
mcb_categories()
table(commodity_costs$category)
```

## Looking up a commodity

```{r}
mcb_costs("copper")[, c("commodity", "cost_basis", "cost_text", "meaningful_metric")]
```

Matching is case-insensitive and falls back to partial matches:

```{r}
mcb_commodities("Precious Metals")
```

## The cost-driver breakdown

Every commodity is decomposed into six driver stages. `mcb_drivers()` returns
this in tidy long form, ready for `dplyr`/`ggplot2`:

```{r}
mcb_drivers("gold")
```

## By-product minerals

A structurally important subset of critical minerals are not mined on their own;
they are recovered as by-products and so have no standalone mining cost:

```{r}
mcb_byproducts()$commodity
```

## Cross-cutting drivers and sources

```{r}
head(mcb_cross_cutting(), 4)
mcb_sources("SOURCES")
```

## India strategic framework

`critical_mineral_strategy` adds policy-oriented attributes for the 30 critical
minerals and joins to `commodity_costs` on the mineral name. The scores are
**indicative starter values, not official figures**.

```{r}
mcb_strategic_groups()
mcb_strategy(min_criticality = 4.5)[, c("official_name", "strategic_group",
                                        "criticality_score", "import_dependence_pct")]
```

## Numeric layer and vintages

For charting, `commodity_costs_numeric` provides parsed `cost_low`/`cost_high`
bounds. Units differ across rows, so filter to one `cost_unit` first:

```{r}
subset(mcb_costs_numeric(category = "Precious Metals"),
       cost_unit == "/oz")[, c("commodity", "cost_low", "cost_high", "cost_unit")]
mcb_vintages()
```

## Caveats and terms of use

Cost figures are **indicative orientation values — not official, audited, or
authoritative numbers** — and are not prices. They can vary 3-5x with grade,
method, depth, geography, currency and by-product credits. For many critical
minerals the **driver logic is more reliable than the dollar figure**. Do not
use the figures as the sole basis for investment, procurement, valuation,
regulatory or policy decisions; always verify against current company guidance
or a commercial cost-curve provider (Wood Mackenzie, CRU, S&P Global) before any
formal use.

**Cross-reference rows.** Copper, Nickel, Tin and PGE (Platinum-Group) appear in
both their base/precious category and the critical category. As of v0.2.0 the
critical-category rows carry their own driver breakdown; a fuller treatment still
lives in the base/precious entry:

```{r}
mcb_drivers("copper", category = "Base Metals")[, c("driver_stage", "driver_text")]
```
